Matroids from families of null hypotheses

### 2013-08-19 presentation/presentation.tex
incorporated willi's changes

Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
\cite{shaffer1986modified}: Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
\begin{itemize}
  \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6$ :
    \begin{itemize}
      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
      \item Dynamic: if, e.g., $H_{\{1,2\}}$ and $H_{\{3,4\}}$ are rejected then only $H_{\{1,3\}}$ and $H_{\{2,3\}}$, or $H_{\{1,4\}}$ and $H_{\{2,3\}}$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$ in this case.
       \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Example 1: 4 Groups}
\input{full4ext}
\end{frame}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{itemize}

\item Consider $\mu_i, i \in \{1,...,m\}$

@@

\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
\begin{itemize}
\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
\item edges correspond to hypotheses, \emph{i.e.}, $E =
      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
      j$, and $e_l \neq e_k$ for $l \neq k$
\end{itemize}
\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
  \in I\}$
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 1: 4 Means}
\input{full4}
\end{frame}

\begin{frame}
\frametitle{General contrast tests (two-sided)}
\begin{itemize}

\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$

\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
  c_{ij} = 0$
\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
  with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
    I}$
\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$

\item Elementary hypotheses are represented by column vectors,
  intersection hypotheses by sub-matrices of the contrast matrix

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 2: Contrast test}

\begin{tabular}{ll}
  \begin{minipage}{.45\textwidth}
    All pairwise comparisons of 4 means 
    \begin{displaymath}
    \bs{C}^t = \left(
      \begin{array}{llll}
      -1 & 1 & 0 & 0 \\
      -1 & 0 & 1 & 0 \\
      -1 & 0 & 0 & 1 \\
      0 & -1 & 1 & 0 \\
      0 & -1 & 0 & 1 \\
      0 & 0 & -1 & 1 
    \end{array}
    \right)
    \end{displaymath}
  \end{minipage} &
  \begin{minipage}{.55\textwidth}
    Modified Williams Contrasts for Dose Response
    \begin{displaymath}
      \bs{C}^t=\left(
      \begin{array}{llll}
        1 & 0 & 0 & -1\\
        1 & 0 &-.51 & -.49 \\
        1 & -.35 & -.33 & -.32\\
        .51 & .49 & -.51 & -.48\\
        .51 & .49 & 0 &  -1\\
        .35 & .33 & .32& -1
      \end{array}\right)
    \end{displaymath}
  \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Dependent null hypotheses}

\begin{itemize}
\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
  dependence relation $\sim_H$ on $\mathcal{P}(S)$
\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq  H_{I\setminus i}$
\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
  column rank
\end{itemize} % ends low level
\end{frame}

\section{Matroids}

\begin{frame}
\frametitle{Matroids}
\begin{itemize}

\item A matroid on $S$ endows $\mathcal{P}(S)$ with an algebraic
  structure 
\item Matroids generalize the structures of graphs and vector spaces,
  \eg circuits, bases, independent sets
\item Many ways to define a matroid
  \begin{itemize}
  \item From a graph perspective: Circuit axioms
  \item From a vector space perspective: Independence axioms
  \end{itemize}
\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms 
\item \textbf{/Some illustration hereabouts/}
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Circuit axioms}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{cycle}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
    \begin{enumerate}
    \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
      $\exists C_3 \subseteq (C_1 \cup C_2)$,
    \end{enumerate}
    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
  \end{minipage}
\end{tabular}

\end{frame}

% \item \textbf{/Some illustration hereabouts/}

% \end{itemize} % ends low level
% \end{frame}

\begin{frame}
\frametitle{Independence axioms}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
  \only<1>{\input{independent}}
  \only<2>{\input{cycle}}
  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
  holds
  \begin{enumerate}
  \item $\emptyset \in \mathscr{I}$,
  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
    \mathscr{I}$,
  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
    \setminus J$ such that $J \cup i \in \mathscr{I}$,
  \end{enumerate}
  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
  on $S$. This defines an independence relation $i \nsim_M I$ iff
  $I\cap i \in \mathscr{I}$.

  \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Dependence axioms}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{dependent}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  Relation $\sim_M$ on elements and subsets of $S$,
  if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
  \begin{enumerate}
  \item $y \sim Y$ for all $y \in Y$
  \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
    then $y \sim x \cup (Y \setminus y)$
  \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
  \end{enumerate}
  then $\sim_M$ defines a matroid on $S$.
  \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}{Maximally independent sets}
  \begin{itemize}
  \item Bases are maximally independent sets of $S$
  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
  \item Circuits are minimally dependent sets, \ie each subset of $C$
    with rank $\rho C - 1$ is a maximally independent set of $C$
  \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Maximally dependent sets}
\begin{itemize}
\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
\item Closed sets {\em flats}: $\mathscr{L} = \{I \subseteq S: \sigma I = I\}$
\item Hyperplanes are flats of rank $\rho S - 1$
\item All flats are given by closing hyperplanes under intersection
\item Circuits in $M^*$ are the complements of hyperplanes in $M$
\end{itemize} % ends low level
\end{frame}

\begin{frame}{Examples}
\framesubtitle<1>{Independent set}  
\framesubtitle<2>{Base}  
\framesubtitle<4>{Dependent set}  
\framesubtitle<3>{Cycle}  
\framesubtitle<5>{Closure}  
\only<1>{\input{independent}}
\only<2>{\input{base}}
\only<4>{\input{dependent}}
\only<3>{\input{cycle}}
\only<5>{\input{closure}}
\end{frame}

\section{Matroids from logically dependent hypotheses}

\begin{frame}
\frametitle{Matroids from logically dependent hypotheses}

  \begin{block}{Theorem}
    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
    S$, an intersection hypotheses $H_I$ is independent iff the
    corresponding edge set $\{e_i:i \in I\}$ is independent in the
    cycle matroid on $\mathscr{G}_I$.

For a family of two-sided
    general contrast tests $H_i$, $i \in S$, an intersection
    hypotheses $H_I$ is independent iff the corresponding contrast
    matrix $\bs{C}_I$ is independent in the vectorial matroid on
    $\mathbb{R}^m$.

    \begin{displaymath}
      i \sim_H I \Leftrightarrow i \sim_M I
    \end{displaymath}
  \end{block}

\end{frame}

\begin{frame}
\frametitle{Closed test procedure}
\begin{itemize}

\item The flats of a matroid are closed under intersection

\item Every elementary hypothesis is a flat

\item Using local level-$\alpha$ tests for each flat we have a closed test procedure that controls the FWER in the strong sense (Closed test principle)

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Examples}
\begin{itemize}

\item Pairwise comparisons

\item The flats of the complete graph correspond to the partitions of
  $\{1,...,m\}$
\item This provides efficient methods to generate all flats for
  pairwise mean comparisons \cite{er1988fast,knuth2005art,kokosinski2006new}

\item The flats of a matroid generate a lattice, procedures testing not all possible comparisons correspond to intervals on that lattice

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Connections to existing procedures}

\textbf{Old wine in new bottles ((c) L. Hothorn)}
\begin{itemize}
\item \cite{westfall1997multiple} checks each intersection hypotheses
  whether it is flat using dependence arguments.
\item \cite{westfall2007multiple} provide a branch and bound algorithm
  to generate all flats. They note that for pairwise comparisons all
  exhaustive hypotheses are {\em partition hypotheses} 
\item \cite{bergmann1988improvements} define the concepts of an
  exhaustive sets
\item \cite{bernhard1991computergestuetzte} shows properties of
  exhaustive sets that are basically {\em closure axioms} of a matroid
  and provides algorithms to check each intersection hypotheses if
  flat using rank function.

  \cite{hommel1999bonferroni} show {\em closure axioms} for exhaustive
  sets 
\item Flats <-> Weakly Exhaustive set (Hommel, Bergmann)

\item Lemma 1.5 in (Bernhard) about weakly exhaustive sets are ``closure axioms'' of a matroid

\item Graphical procedures for pairwise comparisons (Weichert) identify dependence using cycle in the graph

\item \emph{However:}
\begin{itemize}
\item proofs are shorter (more elegant),
\item efficient algorithms exist to find the flats of matroids,
\item we may define procedures using independent sets,
\item \ldots{}
\end{itemize}

\end{itemize} % ends low level
\end{frame}

\begin{frame}{Matroid procedures for pairwise mean comparisons}

\begin{tabular}{ll}
  \begin{minipage}{.3\textwidth}
    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
     {\tiny wikipedia:Bell\_number}
  \end{minipage} &
  \begin{minipage}{.7\textwidth}
  \begin{itemize}
  \item Two-sided comparisons between $m$ means
  \item The cycle matroid of the corresponding graph representation
    provides an equivalent dependence structure
  \item The flats of a (simple) complete graph are
    isomporphic to the partitions of a set of size $m$
  \item Provides an efficient way to generate all exhaustive sets by
    enumeration of all partitions or hyperplanes \cite{er1988fast,knuth2005art,kokosinski2006new} 
  \item The number of flats is given by the Bell-numbers \ie 5, 15,
    52, 203, 877, 4140, ... 
  \end{itemize}
  \end{minipage}
\end{tabular}

\end{frame}

\section{Graphical approaches for logically dependent hypotheses}

\begin{frame}
\frametitle{Reminder: Graphical approaches}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \begin{tikzpicture}[remember picture,overlay]
      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
    \end{tikzpicture}

  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{itemize}
    \item Provides weights $w_{i,I}$ for each elementary hypotheses
      $H_i$ and each non-empty subset $I \subset S$
    \item Weights are used to define a closed test procedure of
      weighted intersection hypotheses tests
    \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
    \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
    \end{itemize}
  \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
  \frametitle{Example: Hierarchical LSD}
\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{lsd2.tex}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{itemize}
    \item 3 Hypotheses:
      \begin{enumerate}
      \item Placebo vs. High Dose
      \item Placebo vs. Low Dose
      \item High Dose vs. Low Dose
      \end{enumerate}
    \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
    \item Test $H_1$ at level $\alpha$
    \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
      simultaneously
    \item Can we generalize this?
    \end{itemize}
  \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
  \frametitle{Fixed sequence test of pairwise mean comparisons}
\begin{itemize}
\item {\bf Picture of hierarchical test hereabouts}
\end{itemize}

\begin{block}{Theorem}
  Assume that $M(\mathscr{H})$ neither contains loops nor
  parallels. Assume that hypotheses $H_1$ through $H_r$ have been
  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are
  to be tested at level $\alpha_r$ according to a prefixed order.  If
  for some $1 \leq s \leq t$ holds
  \begin{equation}
    \label{eq:cond.hierarchical}
    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{(r+l),(r+k)\},
  \end{equation}
  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
  at full level $\alpha_ra$.
\end{block}

\end{frame}

\begin{frame}
  \frametitle{Example: Multiple doses}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \begin{tikzpicture}[overlay]
      \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
    \end{tikzpicture}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{tikzpicture}[remember picture,overlay]
      \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
    \end{tikzpicture}
  \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
  \frametitle{General weighting strategies}
  \framesubtitle{QVA Example}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \begin{tikzpicture}[remember picture,overlay]
      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
    \end{tikzpicture}

  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{tikzpicture}[remember picture,overlay]
      \node[yshift=-50bp,xshift=-180bp] at (current page.north east){\input{qvameans}};
  \end{tikzpicture}
  \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
\frametitle{General weighting strategies}

>>>>

Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
\cite{shaffer1986modified}: Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
\begin{itemize}
  \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6$ :
    \begin{itemize}
      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
      \item Dynamic: $t^*_j \le t_j$, hence more powerful than "static" 
           \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{tabular}{ll}
  \begin{minipage}{.3\textwidth}
    \only<2->{\input{full4}}
.
  \end{minipage} &
  \begin{minipage}{.7\textwidth}
  \begin{itemize}
\item Consider $\mu_i, i \in \{1,...,m\}$

@@

\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
\begin{itemize}
\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
\item edges correspond to hypotheses, \emph{i.e.}, $E =
      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
      j$, and $e_l \neq e_k$ for $l \neq k$
\end{itemize}
\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
  \in I\}$
\end{itemize} % ends low level
  \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
\frametitle{Max number of true hypotheses}
\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
   \input{shaffer}
.
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  \begin{itemize}
  \item "Static":  $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": $t^*_j $ = max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6, j = 2$ :
    \begin{itemize}
      \item  {Static: if, e.g., $H_1$ and $H_2$ are false then at most $H_4, H_5 $ and $H_6$ can be simultaneously true, hence $t_2 = 3$}
      \item  {Dynamic: if, e.g., $H_1$ and $H_6$ are false then only $H_3$ and $H_4$, or $H_3$ and $H_5$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$.}
      \end{itemize}

\end{itemize} % ends low level
  \end{minipage}
\end{tabular}
\end{frame}
%
%
%\begin{frame}
%\frametitle{Pairwise mean comparisons (two-sided)}
%\begin{itemize}
%
%\item Consider $\mu_i, i \in \{1,...,m\}$
%
%\item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$
%
%\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
%\begin{itemize}
%\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
%\item edges correspond to hypotheses, \emph{i.e.}, $E =
%      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
%      j$, and $e_l \neq e_k$ for $l \neq k$
%\end{itemize}
%\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
%\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
%  \in I\}$
%\end{itemize} % ends low level
%\end{frame}
%
%\begin{frame}
%\frametitle{Example 1: 4 Groups}
%\input{full4ext}
%\end{frame}
%

% \begin{frame}
% \frametitle{Example 1: 4 Means}
% \input{full4}
% \end{frame}

%
%\begin{frame}
%\frametitle{General contrast tests (two-sided)}
%\begin{itemize}
%
%\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$
%
%\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
%  c_{ij} = 0$
%\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
%  with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
%    I}$
%\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
%\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$
%
%\item Elementary hypotheses are represented by column vectors,
%  intersection hypotheses by sub-matrices of the contrast matrix
%
%\end{itemize} % ends low level
%\end{frame}
%

%
%\begin{frame}
%\frametitle{Example 2: Contrast test}
%
%
%\begin{tabular}{ll}
%  \begin{minipage}{.45\textwidth}
%    All pairwise comparisons of 4 means
%    \begin{displaymath}
%    \bs{C}^t = \left(
%      \begin{array}{llll}
%      -1 & 1 & 0 & 0 \\
%      -1 & 0 & 1 & 0 \\
%      -1 & 0 & 0 & 1 \\
%      0 & -1 & 1 & 0 \\
%      0 & -1 & 0 & 1 \\
%      0 & 0 & -1 & 1
%    \end{array}
%    \right)
%    \end{displaymath}
%  \end{minipage} &
%  \begin{minipage}{.55\textwidth}
%    Modified Williams Contrasts for Dose Response
%    \begin{displaymath}
%      \bs{C}^t=\left(
%      \begin{array}{llll}
%        1 & 0 & 0 & -1\\
%        1 & 0 &-.51 & -.49 \\
%        1 & -.35 & -.33 & -.32\\
%        .51 & .49 & -.51 & -.48\\
%        .51 & .49 & 0 &  -1\\
%        .35 & .33 & .32& -1
%      \end{array}\right)
%    \end{displaymath}
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Dependent null hypotheses}
%
%\begin{itemize}
%\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
%  dependence relation $\sim_H$ on $\mathcal{P}(S)$
%\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
%\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq
%  H_{I\setminus i}$ for all $i \in I$
%\pause
%\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
%\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
%  column rank
%\end{itemize} % ends low level
%\end{frame}
%
%
%
%\section{Matroids}
%
%
%
%\begin{frame}
%\frametitle{Matroids}
%\begin{itemize}
%
%\item Matroids generalize the structures of graphs and vector spaces,
%  \eg cycles, bases, independent sets
%\item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
%  structure
%\item Cryptomorphism: many ways to define a matroid:
%  \begin{itemize}
%  \item From a graph perspective: Circuit axioms.
%  \item From a vector space perspective: Independence axioms.
%  \end{itemize}
%\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms
%\end{itemize} % ends low level
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Circuit axioms}
%
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%    \only<2->{\input{cycle}}
%.
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
%    \begin{enumerate}
%    \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
%    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
%      $\exists C_3 \subseteq (C_1 \cup C_2)\setminus i$,
%    \end{enumerate}
%    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%% \item \textbf{/Some illustration hereabouts/}
%
%% \end{itemize} % ends low level
%% \end{frame}
%
%
%\begin{frame}
%\frametitle{Independence axioms}
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%  \only<1>{\input{independent}}
%  \only<2>{\input{cycle}}
%  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
%  holds
%  \begin{enumerate}
%  \item $\emptyset \in \mathscr{I}$,
%  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
%    \mathscr{I}$,
%  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
%    \setminus J$ such that $J \cup i \in \mathscr{I}$,
%  \end{enumerate}
%  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
%  on $S$. This defines an independence relation $i \nsim_M I$ iff
%  $I\cap i \in \mathscr{I}$.
%
%  \end{minipage}
%\end{tabular}
%
%
%
%
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Dependence axioms}
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%    \input{dependent}
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%  Relation $\sim_M$ on elements and subsets of $S$,
%  if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
%  \begin{enumerate}
%  \item $y \sim Y$ for all $y \in Y$
%  \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
%    then $y \sim x \cup (Y \setminus y)$
%  \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
%  \end{enumerate}
%  then $\sim_M$ defines a matroid on $S$.
%  \end{minipage}
%\end{tabular}
%
%
%
%\end{frame}
%
%\section{Matroids from logically dependent hypotheses}
%\begin{frame}
%  \frametitle{Dependent hypotheses}
%  \begin{block}{Equivalence of dependence relations $\sim_H = \sim_M$}
%    \begin{itemize}
%    \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
%    \item<2-> Does this mean $\mathscr{H}$ defines a matroid $M$, such that
%      $\sim_H = \sim_M$?
%    \end{itemize}
%  \end{block}
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Matroids from logically dependent hypotheses}
%
%  \begin{block}{Two-sided pariwise comparisons}
%    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
%    S$, an elementary hypotheses $H_j$ depends on intersection
%    hypotheses $H_I$ iff the corresponding edge $e_j$ is dependent on
%    the edge set $\{e_i:i \in I\}$ of the cycle matroid defined by
%    $\mathscr{G}_I$.
%  \end{block}
%
%  \begin{block}{Two-sided linear contrast tests}
%    For a family of two-sided general contrast tests $H_i$, $i \in S$,
%    an elementary hypothesis $H_j$ depends on intersection hypotheses
%    $H_I$ iff the corresponding contrast vector $\bs{c_j}$ is
%    dependent on the column vectors of the matrix $\bs{C}_I$ in the
%    vectorial matroid defined by $\bs{C}$ on $\mathbb{R}^m$.
%  \end{block}
%
%\end{frame}
%
%
%\begin{frame}
%  \frametitle{Matroids from null hypotheses}
%  {\Large What is it good for?\\
%  \invisible<1>{Structure}}
%\end{frame}
%
%\begin{frame}
%\frametitle{Maximally independent sets}
%  \begin{itemize}
%  \item Bases are maximally independent sets of $S$
%  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
%  \item Circuits are minimally dependent sets, \ie each subset of $C$
%    with rank $\rho C - 1$ is a maximally independent set of $C$
%  \end{itemize}
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Maximally dependent sets}
%\begin{itemize}
%\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
%\item Closed sets {\em flats}: $\mathscr{L} = \{\sigma I: I \subseteq S\}$
%\item Hyperplanes are flats of rank $\rho S - 1$
%\item All flats are given by closing hyperplanes under intersection
%\item Circuits in $M^*$ are the complements of hyperplanes in $M$
%\end{itemize} % ends low level
%\end{frame}
%
%
%
%\begin{frame}{Examples}
%\framesubtitle<1>{Independent set}
%\framesubtitle<2>{Base}
%\framesubtitle<4>{Dependent set}
%\framesubtitle<3>{Cycle}
%\framesubtitle<5>{Closure}
%\only<1>{\input{independent}}
%\only<2>{\input{base}}
%\only<4>{\input{dependent}}
%\only<3>{\input{cycle}}
%\only<5>{\input{closure}}
%\end{frame}
%
%
%
%
%
%
%
%
%\begin{frame}
%\frametitle{Flats}
%\begin{itemize}
%
%\item The flats of a matroid are closed under intersection
%
%\item Every elementary hypothesis is a flat
%\item For each $I \subseteq S$, $H_I = H_{\sigma I}$
%\item Define a stepwise procedure on the flats of $M$
%
%
%\end{itemize} % ends low level
%\end{frame}
%
%\begin{frame}
%  \frametitle{Stepwise test procedure}
%  \begin{block}{Multiple testing procedure for logically related hypotheses}
%    Assume that $\mathscr{H}$ with dependence relation $\sim_H$
%    defines a matroid on $M$ on $S$. Then a test procedure that
%    rejects elementary hypotheses $H_i$ if all $H_E$, $i \in E$, $E
%    \in \mathscr{L}$ are rejected using a local level-$\alpha$ test
%    controls the FWER in the strong sense.
%  \end{block}
%
%\end{frame}
%
%\begin{frame}
%  \frametitle{Stepwise multiple testing procedure on the lattice of
%    flats}
%%
%\begin{tabular}{ll}
%  \begin{minipage}{.3\textwidth}
%    \includegraphics<1>[width=1.2\textwidth]{lattice4.pdf}
%    \includegraphics<2>[width=1.2\textwidth]{lattice4accept.pdf}
%    \includegraphics<3>[width=1.2\textwidth]{lattice4reject.pdf}
%    \includegraphics<4>[width=1.2\textwidth]{lattice4rejacc.pdf}\\
%    {\tiny wikipedia:Partition\_of\_a\_Set}
%  \end{minipage} &
%  \begin{minipage}{.7\textwidth}
%  \begin{itemize}
%  \item All flats are given by intersections of hyperplanes
%    $\rightarrow$ knowing all hyperplanes remaining flats are given by
%    intersection of index sets
%  \item Hyperplanes can by found by algorithms that enumerate the
%    circuits of the dual matroid \cite{boros2003algorithms}
%  \item The partial order of flats defined by inclusion defines a
%    geometric lattice
%  \item<2-> Testing only a subset of elementary hypotheses corresponds
%    to intervals
%  \item<3-> Rejection of elementary hypotheses removes intervals
%  \end{itemize}
%
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%\begin{frame}{Matroid procedures for pairwise mean comparisons}
%\begin{tabular}{ll}
%  \begin{minipage}{.3\textwidth}
%    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
%     {\tiny wikipedia:Bell\_number}
%  \end{minipage} &
%  \begin{minipage}{.7\textwidth}
%  \begin{itemize}
%  \item Two-sided comparisons of $m$ means
%  \item The cycle matroid of the graph representation
%    provides equivalent dependence structure
%  \item The flats of a (simple) complete graph are
%    isomporphic to the partitions of a set of size $m$
%  \item Efficient algorithms to generate all exhaustive sets by
%    enumeration of set partitions \cite{er1988fast,knuth2005art,kokosinski2006new}
%  \item The number of flats are $m$-th Bell-Number $-1$ \ie 4, 14,
%    51, 202, 876, 4139, ...
%  \end{itemize}
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%
%
%
%\begin{frame}
%\frametitle{Connections to existing procedures}
%
%\textbf{Old wine in new bottles ((c) L. Hothorn)}
%\begin{itemize}
%\item \cite{shaffer1986modified} represents pairwise comparisons
%  by set partitions
%\item \cite{bergmann1988improvements} define the concepts of an
%  exhaustive sets
%\item \cite{bernhard1991computergestuetzte} shows properties of
%  exhaustive sets that are basically {\em closure axioms} of a matroid
%  and provides algorithms to check each intersection hypotheses if is
%  a flat using rank function.
%\item \cite{westfall1997multiple} generates flats using dependence
%  arguments in vector spaces.
%\item \cite{weichert2000robuste} suggest an MTP based on clique
%  decomposition of the graph, cliques in the graph are flats
%\item \cite{westfall2007multiple} provide a branch and bound algorithm
%  to generate all flats. They note that for pairwise comparisons all
%  exhaustive hypotheses are {\em partition hypotheses}
